Asymmetric relationships and their effects on coexistence

### 2023-09-18 paper/RefereeResponses2.tex

\noindent C3) The only other minor suggestion is that the authors clarify that in Fig 1 they are simply showing that ATAs do occur in nature.  I was not suggesting that the authors include examples pertinent to their question in Figure 1---as they say, readers may not be expecting this. I don't think that the connection between the ATAs in Fig 1 and the rest of the project are that clear, as I already expressed. But I respect the authors wishes (and their other feedback) that it is worthwhile including these. Just a sentence or so at some point explaining that these are not directly related to the ATAs relevant in the remainder of the paper may avoid confusion. \\

\noindent ***\emph{} \\
%%JAS: is it okay to put this is the caption of fig 1?
%%or in main, "Though not directly related to our questions at hand, 
%and for illustrative purposes, Fig1d,e show ...

\noindent \textbf{Comments of referee 2:} \\

\noindent C4) This is a novel contribution highlighting that asymmetric tail associations in correlations between two variables was a very simple aspect not properly considered before but with huge implications for the maintenance of biodiversity. I congratulate the authors for their work. \\

@@

\noindent ***\emph{Thank you for the very positive feedback and previous comments which helped to improve our manuscript!} \\
%DONE

\noindent \textbf{Comments of referee 3:} \\

\noindent C5) I thank the authors for the change and the nice discussion/replies to my questions. I especially like the new figure 2 which greatly helped my understanding of how to remove ATA. \\

\noindent ***\emph{Many thanks for the positive feedback which improved our manuscript to become more understandle to other readers as well!} \\

@@

\noindent \textbf{Comments of the editor:} \\

\noindent C6) Reviewers are enthusiastic about this paper demonstrating the importance of how variable relationships depending on when they are weak or strong influence competitive exclusion and coexistence. \\

\noindent ***\emph{Thank you for pointing out reviewers' positive feedback.} \\

\noindent C7) The legends of figures are rather long and dense. Figures should be self-explanatory. I suggest adding to all of them the sentence at the end: “See Table 1 for used annotations and abbreviations.” \\

@@

\noindent ***\emph{We }
%%So replace any variable explanations with this sentence? I dont feel that every caption needs this. Particularly after addressing next comment. 
%%Sure can add but I don't feel like this replaces explanation of things and just makes caption longer..

\noindent C8) The legend in Fig. 1 is too long and has variables not present in the graphs but mentioned in other sections of the paper. For example, in line 44-45: “The variables (B1, B2) described in the Introduction…”. I suggest revising if this information is needed at all in this Figure. \\

\noindent ***\emph{. } \\
%%Take out parts of caption that mention (B1, B2)?. As long as the main text mentions that we used noise as pictired in fig one that should okay. But double check

\noindent C9) Main concepts should be described in the introduction in a way that could be understood for a broad audience. In this sense, I suggest adding a short non- mathematical definition for storage effects even if you go in more detail in the Theory section in lines 59-60

Page 2, Lines 59-60 “coexistence into the contributions of various mechanisms, with names such as storage effects (Chesson1994). Storage effects refer to….. Storage effects are reviewed conceptually below and defined formally in Theory. \\

### 2023-09-18 paper/RefereeResponses2.tex

%header material for paper
\title{Responses to referees for ``Asymmetric relationships and their effects on coexistence''}
\date{September 2023}

\begin{document}

\noindent \emph{Dear Editors,} \\

\noindent \emph{Thank you for the decision of ``minor revisions''. We have 
fully addressed the referee and editor comments below and in the manuscript.
We hope the paper can now be rapidly formally accepted. } \\

\noindent \emph{We have replied to referee comments below in italic, with each response preceded
by ``***'', to make our responses easier to find.} \\

\noindent \emph{We submit the following pdf files: The main text pdf, including all figures for readability; 
the Supporting Information; a version of the main text with changes tracked; and all figures as 
separate pdfs, for the production process. The document showing 
the changes we made does not include links such as citations and
references to figures, but it does accurately represent the changes made to the text.
We also submit...
%***DAN: all the latex files and so on.
%*%***DAN: make sure the stuff about putting a latex version of the MT at the end of the SM is all ready to go 
%*%in what I submit.
} \\

@@

\noindent \emph{Yours,}

\noindent \emph{Dan Reuman and Jasmin Albert} \\

\noindent \textbf{Comments of referee 1:} \\

\noindent C1) Thank-you to the authors for their response to my comments.\\

@@

\noindent ***\emph{We appreciate your feedback which helped to improve our manuscript; thank you.} \\

\noindent C2) I have two minor suggestions remaining which I think should be straightforward to implement.

First, to consider introducing a discussion of the robustness to more esoteric definitions of the partial-sharp distributions, and adding this briefly to the main text.  Primarily as the authors already suggest in their response letter, as a signpost for future work.\\

\noindent ***\emph{We added the requested material to section 2.3.} \\

@@

\noindent C3) The only other minor suggestion is that the authors clarify that in Fig 1 they are simply showing that ATAs do occur in nature.  I was not suggesting that the authors include examples pertinent to their question in Figure 1---as they say, readers may not be expecting this. I don't think that the connection between the ATAs in Fig 1 and the rest of the project are that clear, as I already expressed. But I respect the authors wishes (and their other feedback) that it is worthwhile including these. Just a sentence or so at some point explaining that these are not directly related to the ATAs relevant in the remainder of the paper may avoid confusion. \\

\noindent ***\emph{We added a sentence, as requested, in the first paragraph of the Introduction.} \\

@@

\noindent C7) The legends of figures are rather long and dense. Figures should be self-explanatory. I suggest adding to all of them the sentence at the end: “See Table 1 for used annotations and abbreviations.” \\

\noindent ***\emph{We did as requested.} \\

\noindent C8) The legend in Fig. 1 is too long and has variables not present in the graphs but mentioned in other sections of the paper. For example, in line 44-45: “The variables (B1, B2) described in the Introduction…”. I suggest revising if this information is needed at all in this Figure. \\

\noindent ***\emph{We removed the variables, as requested. The editor was right, these were not
needed. Apologies for missing this previously.} \\

@@

Page 2, Lines 59-60 “coexistence into the contributions of various mechanisms, with names such as storage effects (Chesson1994). Storage effects refer to….. Storage effects are reviewed conceptually below and defined formally in Theory. \\

\noindent ***\emph{We did as requested.} \\

\noindent C10)  Page 2, Line 60: Start a new paragraph with the sentence “MCT has been applied to several systems…” \\

\noindent ***\emph{This seemed, to us, to be an odd request, as the sentence which followed this one is the 
last in the paragraph, so putting a new paragraph break where the referee suggested would make a very short, two
sentence paragraph, with no clear topic sentence. Furthermore, the purpose of the paragraph this sentence
is a part of is to introduce MCT, a purpose which is definitely served by these two sentences. So, to us,
these sentences seemed to fit better where they were than in a new paragraph on their own. 
So we thought about why the referee may have made such
request and whether we might be able to address the underlying issue. It seemed to us the confusion might have come
from the fact that, whereas the paragraph starts with general MCT, it then transitions to storage effects, and 
then transitions back
to general MCT at the point of the two sentences the referee has asked us to separate. 
Whereas storage effects are part of MCT, so it makes sense to mention them in a paragraph about MCT,
the transtion to storage effects and then back to general MCT might have been confusing. So we moved the
two sentences in question to earlier in the paragraph. We think this is an improvement and should, hopefully,
address the underlying issue as well as streamlining the transition to the following paragraph.}

### 2023-09-18 paper/Paper.Rnw
Dan thinks the referee responses and corresponding changes to the paper are complete

\noindent \textbf{Running head:} Asymmetric relationships and coexistence \\
\noindent \textbf{Key words:} asymmetric tail association, biodiversity, coexistence, competition, copula, modern coexistence theory, niche differentiation, storage effects \\
\noindent \textbf{Type of article:} Letter\\
\noindent \textbf{Words in Abstract:} 150; \textbf{Words in main text:} 4975; \textbf{Words in text boxes:} 0\\
\noindent \textbf{Number of references:} 39; \textbf{Number of figures:} 5; \textbf{Number of tables:} 1; \textbf{Number of text boxes:} 0\\
\noindent \textbf{Corresponding author:}
Daniel C. Reuman,
Department of Ecology and Evolutionary Biology,
University of Kansas,
2101 Constant Ave,
Lawrence, KS 66047;
Phone: 626-560-7084 \\
\noindent \textbf{Author emails:} pimsupaa@uoregon.edu, reuman@ku.edu\\
\noindent \textbf{Statement of Authorship:} The authors jointly carried out all parts of the study.\\
\noindent \textbf{Data accessibility statement:} All codes associated with this study will be released
in a Zenodo repository prior to publication.

@@

%Intro the paper by intro'ing ATAs
At the core of ecology is the study of relationships between organisms and their environments.
Ecologists may ask, for instance, how local environment relates to 
population growth rate. To study relationships, ecologists 
very commonly use metrics such as correlation. But standard measures of the strength 
and direction of association between variables do not tell the whole story 
\citep{Ghosh_2020_AER}.
To illustrate, the three pairs of variables in Fig. \ref{fig:pedagog1}a-c have the same 
Pearson correlation despite striking differences in association structure: Fig. \ref{fig:pedagog1}b shows a 
symmetric type of association, while Fig. \ref{fig:pedagog1}a,c exhibit what has been called 
``asymmetric tail association'' [ATA; \citealp{Ghosh_2020_AER}]. 
Metrics of association other than correlations are also often used (e.g., various forms of 
regression), but other common methods also reveal only a small portion 
of the information content of the relationship \citep{Nelsen2006}.
ATAs and related ideas had previously been seldom considered in ecology, but recent 
insights have been gained by studying how features of associations beyond standard measures 
impact ecological phenomena \citep{deValpine_2014, Anderson_2019, Popovic_2019, Ghosh_2020_AER, Ghosh_2020_Eco,
Ghosh_2020_ecolevol, Ghosh_2021}. For instance, \cite{Ghosh_2020_Eco} argued that extinction risks can become elevated 
when metapopulations are subject to ATAs, e.g., if environmental ``catastrophes'' are more spatially synchronous 
across component populations
than are years of good environment, limiting rescue effects. 
And \cite{Ghosh_2021} showed that ATAs can alter the stability of 
ecosystem functioning. 
%Though not directly related to our questions at hand, and for illustrative purposes,
Fig. \ref{fig:pedagog1}d,e show contrasting examples of ATAs in nature using 
plankton population density time series; many other examples of ATAs in nature have also been explored 
\citep{Ghosh_2020_AER,Ghosh_2020_ecolevol,Ghosh_2021}. To our knowledge, the implications of ATAs for 
competition have not been investigated. Doing so is our goal here.

%Intro to competition, coexistence, MCT, etc.
Competition and coexistence are research topics at the core of modern ecology 
\citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}. The competitive exclusion principle states 
that two species competing for the same limiting resource cannot coexist at constant densities - one will 
instead drive the other extinct \citep{Hardin_1960}.
Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] is a 
framework for understanding how species nevertheless coexist, and biodiversity is maintained, in real ecosystems,
through the niche partitioning effects of environmental variation.
MCT breaks down coexistence into the contributions of various mechanisms, with names such as 
``storage effects'' \citep{Chesson_1994}. 
Storage effects refer to ... 
%effects of [temporal] niche partitioning which allow species to "store" benefits from favorable conditions to persists in unfavorable conditions
 and are reviewed conceptually below and defined formally in Theory.

MCT has been
applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2006, Adler_2010}, 
tropical trees \citep{Usinowicz_2012}, plankton \citep{Caceres_1997, Descamps_2005, Letten2018}, and corals \citep{Alvarez-Noriega_2020}. 
%Narawani_2013
Although MCT is somewhat mathematically complex, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.

@@

How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
correlation unchanged? Fig. \ref{fig:pedagog2} provides an illustration of the steps we describe here. 
We assume, for simplicity, that the cumulative distribution functions (CDFs) of $E_i$ and 
$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, are strictly monotonic and continuous. 
Letting $\varphi$ be the CDF of a standard normal distribution, we consider the random variables
$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$, which are standard normally
distributed, although they are not necessarily jointly bivariate normal (Fig. \ref{fig:pedagog2}a,b).  
Let $\rho_i$ be their covariance. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
$\E(e_i)=\E(c_{i \bs i})=0$ and $\var(e_i)=\var(c_{i \bs i})=1$, and we denote 
$\cov(e_i,c_{i \bs i})$ by $\ddot{\rho}_i$ (Fig. \ref{fig:pedagog2}c). We define 
$E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$ (Fig. \ref{fig:pedagog2}d). 
The value of $\ddot{\rho}_i$ is determined in one of 
four alternative ways: either by setting $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
$(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated.
For those familiar with copula statistics (see, e.g., \citealp{Genest_2007} and \citealp{Ghosh_2020_AER}), 
we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
The choice of $\ddot{\rho}_i$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are correlated to the same extent as 
$E_i$ and $C_{i \bs i}$, according to whichever definition of correlation is preferable. Thus $E_i^\nat$ and 
$C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se}. 
The question of uniqueness of the distribution $(E_i^\nat,C_{i \bs i}^\nat)$ 
is addressed in SI section \ref{SI-SIsect:uniqueness}.
The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created similarly.  
For comparison, Fig. \ref{fig:pedagog2} also shows, for the example explored by that figure, 
the distribution $(E_i^\shp,C_{i \bs i}^\shp)$
defined above (following \cite{Ellner_2016}).

@@

All computations were done in R on a laptop running Ubuntu Linux. 
Complete computer codes for the project are stored at (link to be added).
%\url{https://github.com/jasminalbert/AsymmetricRelationshipsAndCoexistence.git}.

@@

%Summary-type paragraph of main results
We developed and applied a new extension of modern coexistence theory (MCT) that 
quantifies the impact of asymmetric tail associations (ATAs) on species coexistence. 
Building upon a recent, simulation-based approach to MCT \citep{Ellner_2016, Ellner_2019},
we decomposed storage effects into two new mechanisms, the contribution of ATAs to coexistence,
and the contribution of $EC$ correlation \emph{per se}.
Applications of our extended theory to the lottery model and to a laboratory diatom system demonstrated that  
ATA contributions to storage effects and to species coexistence 
can be comparable in magnitude to contributions of previously recognized mechanisms.
Our results add
coexistence to the list of ecological phenomena on which 
ATAs have been shown to have an important impact \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.  
Though storage effects are typically described as the contribution
of $EC$ \emph{covariance} to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019},
our results suggest that storage effects should be redefined as the 
contribution of \emph{associations} between $E$ and $C$, generally, including ATA effects and
correlation \emph{per se}. The prior terminology, \emph{covariance}, is typically construed 
as reflecting the linear relationship between variables, ignoring complexities of the association
such as ATAs, which we now know can be important. ATAs contributions to coexistence were less important than other mechanisms 
in some contexts, but were very important in other contexts, including our empirically
based example (the diatom system).
Our work therefore demonstrates the potential for ATAs to be an important mechanism of coexistence, generally.
Future work should seek to understand precisely when ATAs are or are not important
for coexistence. 
%For example, the robustness of our choice in specifying a bivariate normal partial sharp distribution,
%though a natural and canonical one, to other more esoteric specifications.

@@

\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig1_a-h.pdf}
\caption{Examples motivating our study of the effects of asymmetric tail associations (ATAs) on competition
and coexistence. (a-c) Bivariate random variables with standard normal marginals showing
alternative patterns of association in the left and right parts of the distributions. 
The two variables in (a) are more strongly related to each
other in the left parts of the distributions, termed ``left-tail association;'' those in (b) 
are symmetrically associated; and those in (c) 
are more strongly related to each other in the right parts, termed ``right-tail association''. All three cases have 
the same Pearson correlation, $\rho$, up to sampling variation, so Pearson correlation does not identify these 
very different patterns of association. For (a, c), we used asymmetrically tail associated cases which are perfectly 
correlated in the lower or upper halves of the distributions, though it is also possible to generate asymmetrically 
tail associated noise with imperfect correlations in both halves of the distributions [\citealp{Nelsen2006,Ghosh_2020_AER};
see also (d,e)]. 
%The variables $(B_1,B_2)$ described in the Introduction and Methods were obtained by generating noise as 
%pictured and then transforming the marginals as specified in Methods.
(d,e) Two examples of ATAs found in nature. Axes are population densities of two plankton species, 
\emph{Ceratium fusus} and \emph{Ceratium furca}, sampled in successive years in two of the 15 locations 
in the North Sea considered by \citep{Ghosh_2020_ecolevol}. ATAs were statistically
significant and differed by location, with associations being primarily in the left tails in 
panel d, and primarily in the right tails in panel e (SI section \ref{SI-SIsect:planktonstats}). 
One can see these asymmetries by noting that the open points in panel d, which are in the left halves of the
distributions, are more tightly clustered around the line $y=x$ than are the closed points,
which are in the right halves of the distributions; and \emph{vice versa} for panel e.
(f-h) Lottery model population simulations% using $(B_1,B_2)$ 
obtained by transforming the respectively corresponding variables (a-c; see Introduction).
See Table 1 for used annotations and abbreviations.
\label{fig:pedagog1}}
\end{figure}

>>>>

\noindent \textbf{Running head:} Asymmetric relationships and coexistence \\
\noindent \textbf{Key words:} asymmetric tail association, biodiversity, coexistence, competition, copula, modern coexistence theory, niche differentiation, storage effects \\
\noindent \textbf{Type of article:} Letter\\
\noindent \textbf{Words in Abstract:} 150; \textbf{Words in main text:} 4975; \textbf{Words in text boxes:} 0\\
\noindent \textbf{Number of references:} 39; \textbf{Number of figures:} 5; \textbf{Number of tables:} 1; \textbf{Number of text boxes:} 0\\
\noindent \textbf{Corresponding author:}
Daniel C. Reuman,
Department of Ecology and Evolutionary Biology,
University of Kansas,
2101 Constant Ave,
Lawrence, KS 66047;
Phone: 626-560-7084 \\
\noindent \textbf{Author emails:} pimsupaa@uoregon.edu, reuman@ku.edu\\
\noindent \textbf{Statement of Authorship:} The authors jointly carried out all parts of the study.\\
\noindent \textbf{Data accessibility statement:} All codes associated with this study have been released in a 
public repsitory.

@@

%Intro the paper by intro'ing ATAs
At the core of ecology is the study of relationships between organisms and their environments.
Ecologists may ask, for instance, how local environment relates to 
population growth rate. To study relationships, ecologists 
very commonly use metrics such as correlation. But standard measures of the strength 
and direction of association between variables do not tell the whole story 
\citep{Ghosh_2020_AER}.
To illustrate, the three pairs of variables in Fig. \ref{fig:pedagog1}a-c have the same 
Pearson correlation despite striking differences in association structure: Fig. \ref{fig:pedagog1}b shows a 
symmetric type of association, while Fig. \ref{fig:pedagog1}a,c exhibit what has been called 
``asymmetric tail association'' [ATA; \citealp{Ghosh_2020_AER}]. 
Metrics of association other than correlations are also often used (e.g., various forms of 
regression), but other common methods also reveal only a small portion 
of the information content of the relationship \citep{Nelsen2006}.
ATAs and related ideas had previously been seldom considered in ecology, but recent 
insights have been gained by studying how features of associations beyond standard measures 
impact ecological phenomena \citep{deValpine_2014, Anderson_2019, Popovic_2019, Ghosh_2020_AER, Ghosh_2020_Eco,
Ghosh_2020_ecolevol, Ghosh_2021}. For instance, \cite{Ghosh_2020_Eco} argued that extinction risks can become elevated 
when metapopulations are subject to ATAs, e.g., if environmental ``catastrophes'' are more spatially synchronous 
across component populations
than are years of good environment, limiting rescue effects. 
And \cite{Ghosh_2021} showed that ATAs can alter the stability of 
ecosystem functioning. 
Fig. \ref{fig:pedagog1}d,e show contrasting examples of ATAs using 
plankton population density time series;
these examples are intended only to empirically illustrate what ATAs look like, as they
use data of a different type from what we focus on in this paper.
Many other examples of ATAs in nature have also been explored 
\citep{Ghosh_2020_AER,Ghosh_2020_ecolevol,Ghosh_2021}. To our knowledge, the implications of ATAs for 
competition have not been investigated. Doing so is our goal here.

%Intro to competition, coexistence, MCT, etc.
Competition and coexistence are research topics at the core of modern ecology 
\citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}. The competitive exclusion principle states 
that two species competing for the same limiting resource cannot coexist at constant densities - one will 
instead drive the other extinct \citep{Hardin_1960}.
Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] is a 
framework for understanding how species nevertheless coexist, and biodiversity is maintained, in real ecosystems,
through the niche partitioning effects of environmental variation.
%moved the following two sentences from below
MCT has been
applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2006, Adler_2010}, 
tropical trees \citep{Usinowicz_2012}, plankton \citep{Caceres_1997, Descamps_2005, Letten2018}, and corals \citep{Alvarez-Noriega_2020}. 
Although MCT is somewhat mathematically complex, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
%
MCT breaks down coexistence into the contributions of various mechanisms, with names such as 
``storage effects'' \citep{Chesson_1994}. 
Storage effects, which are reviewed conceptually below and defined formally in Theory, refer to
a type of niche partitioning which allows species to effectively ``store" population growth 
benefits from favorable time periods in order to persist through unfavorable periods.

@@

How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
correlation unchanged? Fig. \ref{fig:pedagog2} provides an illustration of the steps we describe here. 
We assume, for simplicity, that the cumulative distribution functions (CDFs) of $E_i$ and 
$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, are strictly monotonic and continuous. 
Letting $\varphi$ be the CDF of a standard normal distribution, we consider the random variables
$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$, which are standard normally
distributed, although they are not necessarily jointly bivariate normal (Fig. \ref{fig:pedagog2}a,b).  
Let $\rho_i$ be their covariance. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
$\E(e_i)=\E(c_{i \bs i})=0$ and $\var(e_i)=\var(c_{i \bs i})=1$, and we denote 
$\cov(e_i,c_{i \bs i})$ by $\ddot{\rho}_i$ (Fig. \ref{fig:pedagog2}c). We define 
$E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$ (Fig. \ref{fig:pedagog2}d). 
The value of $\ddot{\rho}_i$ is determined in one of 
four alternative ways: either by setting $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
$(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated.
For those familiar with copula statistics (see, e.g., \citealp{Genest_2007} and \citealp{Ghosh_2020_AER}), 
we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
The choice of $\ddot{\rho}_i$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are correlated to the same extent as 
$E_i$ and $C_{i \bs i}$, according to whichever definition of correlation is preferable. Thus $E_i^\nat$ and 
$C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se}. 
The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created similarly.  
For comparison, Fig. \ref{fig:pedagog2} also shows, for the example explored by that figure, 
the distribution $(E_i^\shp,C_{i \bs i}^\shp)$
defined above (following \cite{Ellner_2016}).

The question of uniqueness of the distribution $(E_i^\nat,C_{i \bs i}^\nat)$ 
is addressed fully in SI section \ref{SI-SIsect:uniqueness}.
Briefly, a choice 
is made as to what type of correlation (e.g., Pearson, Spearman, etc.) is to be preserved. 
Additionally, another choice involves the use of a bivariate normal distribution $(e_i,c_{i \bs i})$; 
any other symmetrically tail associated bivariate 
distribution with standard normal marginals could be used instead.
But the normal distribution is a natural choice, and other choices may 
be somewhat esoteric (SI section \ref{SI-SIsect:uniqueness}). 
Future work could explore the consequences of this choice.

@@

All computations were done in R on a laptop running Ubuntu Linux. 
Complete computer codes for the project are stored at 
\url{https://github.com/jasminalbert/AsymmetricRelationshipsAndCoexistence.git}.

@@

%Summary-type paragraph of main results
We developed and applied a new extension of modern coexistence theory (MCT) that 
quantifies the impact of asymmetric tail associations (ATAs) on species coexistence. 
Building upon a recent, simulation-based approach to MCT \citep{Ellner_2016, Ellner_2019},
we decomposed storage effects into two new mechanisms, the contribution of ATAs to coexistence,
and the contribution of $EC$ correlation \emph{per se}.
Applications of our extended theory to the lottery model and to a laboratory diatom system demonstrated that  
ATA contributions to storage effects and to species coexistence 
can be comparable in magnitude to contributions of previously recognized mechanisms.
Our results add
coexistence to the list of ecological phenomena on which 
ATAs have been shown to have an important impact \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.  
Though storage effects are typically described as the contribution
of $EC$ \emph{covariance} to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019},
our results suggest that storage effects should be redefined as the 
contribution of \emph{associations} between $E$ and $C$, generally, including ATA effects and
correlation \emph{per se}. The prior terminology, \emph{covariance}, is typically construed 
as reflecting the linear relationship between variables, ignoring complexities of the association
such as ATAs, which we now know can be important. ATAs contributions to coexistence were less important than other mechanisms 
in some contexts, but were very important in other contexts, including our empirically
based example (the diatom system).
Our work therefore demonstrates the potential for ATAs to be an important mechanism of coexistence, generally.
Future work should seek to understand precisely when ATAs are or are not important
for coexistence.

@@

\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig1_a-h.pdf}
\caption{Examples motivating our study of the effects of asymmetric tail associations (ATAs) on competition
and coexistence. (a-c) Bivariate random variables with standard normal marginals showing
alternative patterns of association in the left and right parts of the distributions. 
The two variables in (a) are more strongly related to each
other in the left parts of the distributions, termed ``left-tail association;'' those in (b) 
are symmetrically associated; and those in (c) 
are more strongly related to each other in the right parts, termed ``right-tail association''. All three cases have 
the same Pearson correlation, $\rho$, up to sampling variation, so Pearson correlation does not identify these 
very different patterns of association. For (a, c), we used asymmetrically tail associated cases which are perfectly 
correlated in the lower or upper halves of the distributions, though it is also possible to generate asymmetrically 
tail associated noise with imperfect correlations in both halves of the distributions [\citealp{Nelsen2006,Ghosh_2020_AER};
see also (d,e)]. 
%The variables $(B_1,B_2)$ described in the Introduction and Methods were obtained by generating noise as 
%pictured and then transforming the marginals as specified in Methods.
(d,e) Two examples of ATAs found in nature. Axes are population densities of two plankton species, 
\emph{Ceratium fusus} and \emph{Ceratium furca}, sampled in successive years in two of the 15 locations 
in the North Sea considered by \citep{Ghosh_2020_ecolevol}. ATAs were statistically
significant and differed by location, with associations being primarily in the left tails in 
panel d, and primarily in the right tails in panel e (SI section \ref{SI-SIsect:planktonstats}). 
One can see these asymmetries by noting that the open points in panel d, which are in the left halves of the
distributions, are more tightly clustered around the line $y=x$ than are the closed points,
which are in the right halves of the distributions; and \emph{vice versa} for panel e.
(f-h) Lottery model population simulations% using $(B_1,B_2)$ 
obtained by transforming the respectively corresponding variables (a-c; see Introduction).
\label{fig:pedagog1}}
\end{figure}
